### 2015-01-28 Projects/DelCo_2012/Report/DelCo_2012.tex
DelCo 2012: Added Uncertainties

The Engineering Laboratory at National Institute of Standards and Technology has a research program aimed at improving the safety and effectiveness of firefighters through improved knowledge of fire behavior and firefighting tactics.  One of the objectives of this program is to examine non-traditional means of fire suppression.   This series of experiments conducted by NIST for the California Polytechnic State University, San Luis Obispo in collaboration with the Montgomery County (MD) Fire and Rescue Service complements the NIST research program on firefighting tactics.  These experiments are part of a larger research project led by Cal Poly with the purpose of investigating the capabilities and limitations of compressed air foam systems (CAFS) for structural firefighting with the objective of enhancing the scientific knowledge base regarding the effectiveness and safety implications associated with the use of CAFS for structural firefighting.  
NIST conducted several series of experiments to provide scientific measurements and comparisons between CAFS and water-only hose streams for different compartment fire scenarios.  
Three types of experiments were conducted; 
1)            spray density measures under non-fire conditions 
2)            gas layer cooling 
3)            fire suppression in an experimental structures

@@

>>>>

The Engineering Laboratory at National Institute of Standards and Technology has a research program aimed at improving the safety and effectiveness of firefighters through improved knowledge of fire behavior and firefighting tactics.  One of the objectives of this program is to examine non-traditional means of fire suppression.   This series of experiments conducted by NIST for the California Polytechnic State University, San Luis Obispo in collaboration with the Montgomery County (MD) Fire and Rescue Service complements the NIST research program on firefighting tactics.  These experiments are part of a larger research project led by Cal Poly with the purpose of investigating the capabilities and limitations of compressed air foam systems (CAFS) for structural firefighting with the objective of enhancing the scientific knowledge base regarding the effectiveness and safety implications associated with the use of CAFS for structural firefighting.  
NIST conducted several series of experiments to provide scientific measurements and comparisons between CAFS and water-only hose streams for different compartment fire scenarios.

Three types of experiments were conducted;

1)            spray density measures under non-fire conditions 
2)            gas layer cooling 
3)            fire suppression in an experimental structures

@@

There are different components of uncertainty in the length, mass, temperature, heat flux, gas concentration, differential pressure, gas velocity and heat release rate reported here. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means  [ ]. Type B analysis of systematic uncertainties involves estimating the upper (+ a) and lower (- a) limits for the quantity in question such that the probability that the value would be in the interval (± a) is essentially 100 %. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95 % confidence interval (2σ).  For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

Each length measurement was taken carefully. Length measurements such as the room dimensions, instrumentation array locations and fire apparatus (for example nozzle, sprinkler, or fan) placement were made with a hand held laser measurement device which is has an accuracy of ± 6.0 mm (0.25) over a range of 0.61 m (2.00 ft) to 15.3 m (50.0 ft) [ ] .  However, conditions affecting the measurement, such as levelness of the device, yields an estimated uncertainty of ± 0.5 % for measurements in the 2.0 m (6.6 ft) to 10.0 m (32.8 ft) range.  Steel measuring tapes with a resolution of  ± 0.5 mm (0.02 in) were used to locate individual sensors within a measurement array and to measure and position the furniture. The steel measuring tapes were manufactured in compliance with NIST Manual 44, which specifies a tolerance of ±1.6 mm (0.06 in) for 9.1 m (30 ft) tapes and ±6.4 mm (0.25 in) for 30.5 m (100 ft) tapes [ ] .  Some issues, such as “soft” edges on the upholstered furniture, result in an estimated total expanded uncertainty of  ± 1.0 %.

The load cell used to weigh the fuels prior to the experiments had a range of 0 kg (0 lbs) to 200 kg (440lbs) with a resolution of a 0.05 kg (0.11 lb) and a calibration uncertainty within 1% [] . The expanded uncertainty is estimated to be less than  ± 5 %. 
The standard uncertainty in temperature of the thermocouple wire itself is  ± 2.2 °C at 277 °C and increases to ± 9.5 °C at 871 °C as determined by the wire manufacturer [ ] .  The variation of the temperature in the environment surrounding the thermocouple is known to be much greater than that of the wire uncertainty [ ,  ] .  Small diameter thermocouples were used to limit the impact of radiative heating and cooling.  The estimated total expanded uncertainty for temperature in these experiments is ± 15 %.
In this study, total heat flux measurements were made with water-cooled Schimidt-Bolter gauges.  The manufacturer reports a ± 3 % calibration expanded uncertainty for these devices [ ] . Results from an international study on total heat flux gauge calibration and response demonstrated that the uncertainty of a Schmidt-Boelter gauge is typically ± 8 % [] .

The gas measurement instruments and sampling system used in this series of experiments have been demonstrated an expanded (k = 2) relative uncertainty of ± 1 % when compared with span gas volume fractions [].   Given the non-uniformities and movement of the fire gas environment and the limited set of sampling points in these experiments an estimated uncertainty of ± 12 % is being applied to the results [ ].

Differential pressure reading uncertainty components were derived from pressure transducer
instrument specifications and previous experience with pressure transducers. The transducers
were factory calibrated and the zero and span of each was checked in the laboratory prior to the
experiments yielding an accuracy of ± 1 % [ ] . The total expanded uncertainty was estimated at 10 %.

Bi-directional probes and single thermocouples were used to measure the velocity.  The bi-directional probes used similar pressure transducers as those used for the differential pressure measurements discussed above.   Bare-bead Type K thermocouple are co-located with the probe. A gas velocity measurement study, examining the doorway flow of pre-flashover compartment fires, yielded expanded uncertainty measurements ranging from ± 0.14 to ± 0.22 for bi-directional probes of similar design[] .  The total expanded uncertainty for gas velocity in these experiments w estimated to be  ± 18 %.   
Water Flowrate
Water flowrate was measured with a pressure and flow meter combination shown in Figure x.x. The meter consists of a section of 6.35 cm (2.5 inch) cast aluminum pipe with a 0 – 4.1 MPa (0 - 600 psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800 lpm (1250 gpm). The pressure transducer and paddlewheel both connect to the battery operated control box where the pressure transducer voltage is converted to a pressure and the paddlewheel pulse count is converted to a volumetric flow rate.  The manufacturer reports a ± 5 % calibration expanded uncertainty for the flow sensor and ± 3 %  for the pressure sensor [ ] . The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flowrate. The total expanded uncertainty was estimated at ± 10 %.
